---
title: "Practical Machine Learning - Course Project"
author: "Pascal Bellerose"
date: "22 juillet 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose
The goal of this analysis is to predict the manner in which a specific exercise was executed.  5 different possible outcomes:  

* Class A: Executed according to specifications
* Class B: Throwing the elbows to the front
* Class C: Lifting dumbell only halfway
* Class D: Lowering the dumbell only halfway
* Class E: Throwing the hips to the front

# Context Description
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: [link](http://groupware.les.inf.puc-rio.br/har) (see the section on the Weight Lifting Exercise Dataset).  
Participants were asked to perform one set of 10 repetitions
of the Unilateral Dumbbell Biceps Curl in five different fashions:
exactly according to the specification (Class A), throwing
the elbows to the front (Class B), lifting the dumbbell
only halfway (Class C), lowering the dumbbell only halfway
(Class D) and throwing the hips to the front (Class E). Class
A corresponds to the specified execution of the exercise,
while the other 4 classes correspond to common mistakes.

# Load and process data

```{r}
  library(data.table)
  library(caret)
  library(randomForest)
  library(AppliedPredictiveModeling)
  pml.testing = read.csv("pml-testing.csv", row.names = 1, na.strings = c("NA","#DIV/0!"))
  pml.training = read.csv("pml-training.csv", row.names = 1, na.strings = c("NA","#DIV/0!"))
```


## Preprocessing

  
#### Exclude near Zero Variance Predictors
```{r nearZeroVar}
  excluded <- nearZeroVar(pml.training)
  training <- pml.training[, -excluded]
  dim(training)
```
#### Exclude predictors having more than 80% of NAs
```{r naRemove}
  training <- training[ , !colMeans(is.na(training))>.8]
  dim(training)
```

#### Exclude descriptors
```{r descrRemove}
  training <- training[, -c(1:5)]
  dim(training)
```


## Correlated Predictors
```{r}
  corPreds <- cor(x=training[,-53])
  summary(corPreds[upper.tri(corPreds)])
```
  
Since there are some highly correlated predictors, it might be better to reduce the impact of those correlations.  
  
```{r}
  highCor <- findCorrelation(corPreds)
  training <- training[,-highCor]
  dim(training)
```

Since there is a lot of variance in the predictors
```{r}
  inTrain = createDataPartition(y=training$classe, p=0.80, list=FALSE)
  t1 = training[inTrain,]
  v1 = training[-inTrain,]
  library(doParallel)
  cl = makeCluster(detectCores()-2)
  registerDoParallel(cl)
  set.seed(1017)
  fit = train(classe ~ ., data=t1, 
              method='rf')
  pred = predict(fit, v1)
  ##stopCluster(cl)
  predAcc = confusionMatrix(pred, v1$classe)
  fit
```




